---
title: "Getting Familiar with Apollo: RP"
output: html_document
date: "2024-02-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MNL Revealed Preferences (Multinomial Logit model)

Our first resource is a synthetic dataset looking at mode choice for 500 travellers. For each individual, the data contains two revealed preference (RP) inter-city trips, where the possible modes were **car**, **bus**, **air** and **rail**, and where each individual has at least two of these four modes available to them. The journey options are described on the basis of access time (except for car), travel time and cost, with times in minutes, and costs in $. The data then also contains 14 stated preference (SP) tasks per person, using the same alternatives as those available on the RP journey for that person, but with an additional categorical quality of service attribute added in for air and rail, taking three levels, namely *no frills*, *wifi available*, or *food available*. For each individual, the dataset also contains information on gender, whether the journey was a business trip or not, and the individual’s income.

### Library and settings

```{r library and settings}
### Clear memory
rm(list = ls())

### Load Apollo library
library(apollo)

### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_RP",
  modelDescr      = "Simple MNL model on mode choice RP data",
  indivID         = "ID", #indicate which columns of the data is the decision maker
  outputDirectory = "output"
)
```
### Loading data from package

```{r Loading data}

### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)

database = apollo_modeChoiceData #simulated dataset, 8'000 rows, 26 variables (25 + id)
database #long format
colnames(database) #see column names

### for data dictionary, use ?apollo_modeChoiceData

### Use only RP data
database = subset(database,database$RP==1)
```


### Define Model Parameters

```{r}
### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(asc_car   = 0,
              asc_bus   = 0,
              asc_air   = 0,
              asc_rail  = 0,
              b_tt_car  = 0,
              b_tt_bus  = 0,
              b_tt_air  = 0,
              b_tt_rail = 0,
              b_access  = 0,
              b_cost    = 0)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none

apollo_fixed = c("asc_car") #ASC that will not be estimated, should keep it empty (apollo_fixed = c() if all param need to be estimated)

```

### Group and validate input, define model + likehood function                      

```{r}


apollo_inputs = apollo_validateInputs() #Searches the user work space (.GlobalEnv) for all necessary input to run apollo_estimate, and packs it in a single list.


apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){
  
  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  
  ### Create list of probabilities P
  P = list()
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  V[["car"]]  = asc_car  + b_tt_car  * time_car                           + b_cost * cost_car
  V[["bus"]]  = asc_bus  + b_tt_bus  * time_bus  + b_access * access_bus  + b_cost * cost_bus 
  V[["air"]]  = asc_air  + b_tt_air  * time_air  + b_access * access_air  + b_cost * cost_air   
  V[["rail"]] = asc_rail + b_tt_rail * time_rail + b_access * access_rail + b_cost * cost_rail  
  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(car=1, bus=2, air=3, rail=4), 
    avail         = list(car=av_car, bus=av_bus, air=av_air, rail=av_rail), 
    choiceVar     = choice,
    utilities     = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

```
### Model Estimation and output

```{r}
model = apollo_estimate(apollo_beta, apollo_fixed, apollo_probabilities, apollo_inputs)


apollo_modelOutput(model)


apollo_saveOutput(model)

```
