---
title: "Test_Apollo_SP_effects"
output: html_document
date: "2024-02-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ################################################################# #
#### LOAD LIBRARY AND DEFINE CORE SETTINGS                       ####
# ################################################################# #

### Clear memory
rm(list = ls())





```{r library and settings}
### Clear memory
rm(list = ls())

### Load Apollo library
library(apollo)

### Initialise code
apollo_initialise()

### Set core controls
apollo_control = list(
  modelName       = "MNL_SP_effects",
  modelDescr      = "Simple MNL model on mode choice SP data using effects coding",
  indivID         = "ID",
  outputDirectory = "output"
)
```
### Loading data from package

```{r Loading data}

### Loading data from package
### if data is to be loaded from a file (e.g. called data.csv), 
### the code would be: database = read.csv("data.csv",header=TRUE)
database = apollo_modeChoiceData
### for data dictionary, use ?apollo_modeChoiceData

### Use only SP data
database = subset(database,database$SP==1)
```

### Define Model Parameters

```{r}
### Vector of parameters, including any that are kept fixed in estimation
apollo_beta=c(asc_car      = 0,
              asc_bus      = 0,
              asc_air      = 0,
              asc_rail     = 0,
              b_tt_car     = 0,
              b_tt_bus     = 0,
              b_tt_air     = 0,
              b_tt_rail    = 0,
              b_access     = 0,
              b_cost       = 0,
              b_wifi       = 0,
              b_food       = 0)

### Vector with names (in quotes) of parameters to be kept fixed at their starting value in apollo_beta, use apollo_beta_fixed = c() if none
apollo_fixed = c("asc_car")
```

### Group and validate input, define model + likehood function                      

```{r}


apollo_inputs = apollo_validateInputs()


apollo_probabilities=function(apollo_beta, apollo_inputs, functionality="estimate"){
    
  ### Attach inputs and detach after function exit
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  ### Create list of probabilities P
  P = list()
  
  ### Effects coding constraint
  b_no_frills = - b_wifi - b_food
  
  ### List of utilities: these must use the same names as in mnl_settings, order is irrelevant
  V = list()
  V[["car"]]  = asc_car  + b_tt_car  * time_car                           + b_cost * cost_car
  V[["bus"]]  = asc_bus  + b_tt_bus  * time_bus  + b_access * access_bus  + b_cost * cost_bus 
  V[["air"]]  = asc_air  + b_tt_air  * time_air  + b_access * access_air  + b_cost * cost_air    + b_no_frills * ( service_air == 1 )  + b_wifi * ( service_air == 2 )  + b_food * ( service_air == 3 )
  V[["rail"]] = asc_rail + b_tt_rail * time_rail + b_access * access_rail + b_cost * cost_rail   + b_no_frills * ( service_rail == 1 ) + b_wifi * ( service_rail == 2 ) + b_food * ( service_rail == 3 )
  
  ### Define settings for MNL model component
  mnl_settings = list(
    alternatives  = c(car=1, bus=2, air=3, rail=4), 
    avail         = list(car=av_car, bus=av_bus, air=av_air, rail=av_rail), 
    choiceVar     = choice,
    utilities     = V
  )
  
  ### Compute probabilities using MNL model
  P[["model"]] = apollo_mnl(mnl_settings, functionality)
  
  ### Take product across observation for same individual
  P = apollo_panelProd(P, apollo_inputs, functionality)
  
  ### Prepare and return outputs of function
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

```

# ################################################################# #
#### MODEL ESTIMATION                                            ####
# ################################################################# #

model = apollo_estimate(apollo_beta, apollo_fixed, 
                        apollo_probabilities, apollo_inputs)

# ################################################################# #
#### MODEL OUTPUTS                                               ####
# ################################################################# #

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO SCREEN)                               ----
# ----------------------------------------------------------------- #

apollo_modelOutput(model)

# ----------------------------------------------------------------- #
#---- FORMATTED OUTPUT (TO FILE, using model name)               ----
# ----------------------------------------------------------------- #

apollo_saveOutput(model)

# ################################################################# #
##### POST-PROCESSING                                            ####
# ################################################################# #

### Print outputs of additional diagnostics to new output file (remember to close file writing when complete)
apollo_sink()

### calculate value and standard error for base of effects coded parameter
apollo_deltaMethod(model,deltaMethod_settings = list(expression=c(b_no_frills="-b_wifi-b_food")))

# ----------------------------------------------------------------- #
#---- switch off writing to file                                 ----
# ----------------------------------------------------------------- #

apollo_sink()
